---
title: "Visualization I"
author: "Nicole Criscuolo"
date: "2025-09-25"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(ggridges)
library(p8105.datasets)
```
 
Import weather data.
```{r}
data("weather_df")
```

Making first plot (scatter plot of min temps to max temps) (plus sign is to ggplot as pipes are to tidy data calls)
```{r}
ggplot(data = weather_df, mapping = aes(x = tmin, y = tmax)) +
  geom_point()
```

Same thing as above
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point()
```

Also same thing as above
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

```{r}
ggp_weather_scatterplot =
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()

ggp_weather_scatterplot
```

How to check some rows are missing
```{r}
weather_df |> 
  filter(is.na(tmax))
```

## Fancier scatterplots

`alpha` is for opacity ranging from 0-1, closer to 1 is more opaque
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .3) +
  geom_smooth(se = FALSE)
```

Where you define aesthetics can matter.
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .3, size = .6) +
  geom_smooth(se = FALSE)
```

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name, size = prcp), alpha = .3) +
  geom_smooth(se = FALSE)
```

Use faceting (breaks single plot into multiple plots)
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .3, size = 0.8) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .3, size = 0.8) +
  geom_smooth(se = FALSE) +
  facet_grid(name ~ .)
```

More interesting plot
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name, shape = name)) +
  geom_point(aes(size = prcp), alpha = .3) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```

Write a code chain that starts with weather_df; focuses only on Central Park, converts temperatures to Fahrenheit, makes a scatterplot of min vs. max temperature, and overlays a linear regression line (using options in geom_smooth()). ?geom_smooth

```{r}
weather_df |> 
  filter(name == "CentralPark_NY") |> 
  mutate(
    tmax_fahr = tmax * (9/5) + 32,
    tmin_fahr = tmin * (9/5) + 32
  ) |> 
  ggplot(aes(x = tmin_fahr, y = tmax_fahr)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```




